---
title: "air_quality_inequity"
author: "Odera Nwosu"
date: "2023-12-13"
output: html_document
---

```{r}
library(lubridate)
library(httr)
```

## Reading In & Formatting Data

::: {#formatting}
## Reading In & Formatting Data

```{r}
response <- VERB("GET", 'https://api.openaq.org/v2/measurements?location_id=665&parameter=pm25&date_from=2020-12-11T19:00:00-05:00&date_to=2023-12-13T13:43:09-05:00&limit=30000', accept("application/json"))

bronx_aq <- jsonlite::fromJSON( content(response, "text") )$results

response <- VERB("GET", 'https://api.openaq.org/v2/measurements?location_id=228663&parameter=pm25&date_from=2020-12-11T19:00:00-05:00&date_to=2023-12-13T13:53:27-05:00&limit=30000', accept("application/json"))

williamsburg_aq <- jsonlite::fromJSON( content(response, "text") )$results
```

```{r}
```

```{r}
# data with timestamps
modify_dates <- function(data) {
  data$timestamp <- ymd_hms(data$date$utc, tz = "UTC")
  return(data)
}
bronx_aq <- modify_dates(bronx_aq)
williamsburg_aq <- modify_dates(williamsburg_aq)
start_date <- as.POSIXct("2023-01-01 03:00:00")
end_date <- min(c(max(bronx_aq$timestamp), max(williamsburg_aq$timestamp)))

bronx_aq <- bronx_aq[bronx_aq$timestamp > start_date & bronx_aq$timestamp < end_date, ]
williamsburg_aq <- williamsburg_aq[williamsburg_aq$timestamp > start_date & williamsburg_aq$timestamp < end_date, ]

timeline_string <- paste("from", start_date, "to", end_date)
```
:::

::: {#raw_analysis}
## Raw Analysis of Particulate Matter (pm25) Measurements

### I. Distribution of Particulate Matter Data in Bronx, NY

```{r}
hist(bronx_aq$value, main = paste("Particulate Matter Measurements in Bronx NY", timeline_string), cex.main=.8, xlab = "Air Quality Index (AQI)", col="skyblue", breaks = 100)
```

```{r}
boxplot(data$pm25, names = c("Particulate Matter (pm25)"), ylab="Aiq Quality Index", main = "Box Plot Comparison of pm25", col = c("skyblue"))


```

### II. Distribution of Particulate Matter Data in Williamsburg, NY

```{r}
hist(williamsburg_aq$value, main = paste("Particulate Matter (pm25 µg/m³) Measurements in Bronx NY", timeline_string), cex.main=.8, xlab = "Air Quality Index (AQI)", col="salmon1", breaks = 100)

```

### III. Box plot Comparison of Data

```{r}
boxplot(bronx_aq$value, williamsburg_aq$value, names = c("The Bronx, NYC", "Lower Williamsburg, NYC"), ylab="Aiq Quality Index", main = "Box Plot Comparison between pm25 and o3", col = c("skyblue", "salmon1"))
```

## Paired T-Difference Procedure

### Formatting Data

```{r}
merged_aq_data <- merge(bronx_aq, williamsburg_aq, by = "timestamp", all = FALSE)
## If the difference is negative, then Lower williamsburg had a larger pm25 measure.
merged_aq_data$difference <- merged_aq_data$value.x - merged_aq_data$value.y
```
:::

::: {#one-sample-t-test}
## One Sample T-Test

### Hypothesis

```{r}
null_hypothesis_difference <- 0
alpha_level <- 0.5
```

### Checking Assumptions

#### Random Sampling

```{r}
all_days <- seq(start_date, end_date, by = "hour")
missing_differences_days <- setdiff(all_days, merged_aq_data$timestamp)
```

```{r}
plot(all_days, rep(1, length(all_days)), type = "n", xlab = "Time", ylab = "", main = paste("Missing Hours from Paired Data between", start_date,"to", end_date), cex.main =.7)
points(missing_differences_days, rep(1, length(missing_differences_days)), cex = 0.7, pch = 4, col = "red")
```

#### Normal Distribution

```{r}
hist(merged_aq_data$difference, main = paste("Difference between pm25 measures from The Bronx and williamsburg", timeline_string), cex.main=.7, xlab = "Air Quality Index (AQI)", col="skyblue", breaks = 100)
```

```{r}
length(merged_aq_data$difference) > 30
```

#### Independence

### Calculations

```{r}
x <- mean(merged_aq_data$difference)
n <- length(merged_aq_data$difference)
df <- n - 1
s <- sd(merged_aq_data$difference)

```

```{r}
se <- s/sqrt(n)
cat("standard error:", se, "\n")
t_score <- (x - null_hypothesis_difference)/se
cat("t-score:", t_score, "\n")
```

```{r}
p_stat <- pt(t_score, df=df)
cat("p-stat:", p_stat, "\n")
p_value <- 1 - p_stat
cat("p-value:", p_value)
```

```{r}

```
:::
